---
title: "JAN20 DATA ASSESSMENT"
author: "Chao"
date: "3/22/2020"
output: 
  html_notebook:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reshape2)
library(ggplot2)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(configr)
library(tools)
library(stringi)
md5 <- function(f,m){
  if(tools::md5sum(f)==m){
    warning("MD5 check passed.\n")
    return(f)
  }else{
    stop("MD5 check failed. Please check the version and content!\n")
  }
}
```

# Background

Samples involved:

| ID | lane | barcode | Kindom | RCA | Location    |
| -- | ---- | ------- | -------| --- | ----------- |
| U1 | FF1  | 1,17    | Bac    | RCA | Urine       |
| S1 | FF1  | 3,4     | Bac    | RCA | Soil spot A |
| O1 | FF1  | 5,22    | Bac    | RCA | Soil spot B |
| M1 | FF1  | 7       | Bac    | RCA | Mock        |
| Z1 | FF1  | 8       | Bac    | RCA | ZYMO 8      |

Zymo mock strains:

| Species                  | Genomic DNA | 16S Only1 | 16S & 18S1 | Genome Copy2 | Cell Number3 |
| ------------------------ | ----------- | --------- | ---------- | ------------ | ------------ |
| Pseudomonas aeruginosa   | 12          | 4.2       | 3.6        | 6.1          | 6.1          | +
| Escherichia coli         | 12          | 10.1      | 8.9        | 8.5          | 8.5          | +
| Salmonella enterica      | 12          | 10.4      | 9.1        | 8.7          | 8.8          | +
| Lactobacillus fermentum  | 12          | 18.4      | 16.1       | 21.6         | 21.9         |
| Enterococcus faecalis    | 12          | 9.9       | 8.7        | 14.6         | 14.6         | +
| Staphylococcus aureus    | 12          | 15.5      | 13.6       | 15.2         | 15.3         | +
| Listeria monocytogenes   | 12          | 14.1      | 12.4       | 13.9         | 13.9         | +
| Bacillus subtilis        | 12          | 17.4      | 15.3       | 10.3         | 10.3         | +
| Saccharomyces cerevisiae | 2           | NA        | 9.3        | 0.57         | 0.29         |
| Cryptococcus neoformans  | 2           | NA        | 3.3        | 0.37         | 0.18         |

#stat blast annotation of clips fasta to silva database
```{bash,eval=FALSE}
mkdir -p ../../Results/JAN20/STAT/mash
mkdir -p ../../Results/JAN20/STAT/CLIP

# bac
metabbq smk -j -npk SAM/Z{1,2,3}/CLIP/clip2CloseRef.bead.anno
for i in Z{1,2,3};do awk -v ID=$i '{print ID"\t"$0}' SAM/$i/CLIP/clip2CloseRef.bead.anno;
done > ../../Results/JAN20/STAT/CLIP/B.M.clip2CloseRef.bead.anno

for i in Z{1,2,3};do awk -v ID=$i '{print ID"\t"$0}' SAM/$i/CLIP/clip2CloseRef.ref.cov;
done > ../../Results/JAN20/STAT/CLIP/B.M.clip2CloseRef.ref.cov

for i in Z{1,2,3};do 　awk -v i=$i '{b=sprintf("BI%08d",FNR);print i"\t"b"\t"$0}' SAM/$i/mash/BI.msh.tsv 
done > ../../Results/JAN20/STAT/mash/BI.msh.tsv

# fungi
metabbq smk --configfile F.config.yaml -j -npk SAM/M{4,5,6}/CLIP/clip2CloseRef.bead.anno
for i in M{4,5,6};do awk -v ID=$i '{print ID"\t"$0}' SAM/$i/CLIP/clip2CloseRef.bead.anno;
done > ../../Results/JAN20/STAT/CLIP/F.M.clip2CloseRef.bead.anno

for i in M{4,5,6};do awk -v ID=$i '{print ID"\t"$0}' SAM/$i/CLIP/clip2CloseRef.ref.cov;
done > ../../Results/JAN20/STAT/CLIP/F.M.clip2CloseRef.ref.cov

for i in M{4,5,6};do awk -v i=$i '{b=sprintf("BI%08d",FNR);print i"\t"b"\t"$0}' SAM/$i/mash/BI.msh.tsv;
done > ../../Results/JAN20/STAT/mash/F.M.BI.msh.tsv

#test
mkdir -p ../../Results/JAN20/STAT/READ
for i in Z{1,7};do awk -v ID=$i '{print ID"\t"$0}' SAM/$i/READ/fastp2CloseRef.bam.depth;
done > ../../Results/JAN20/STAT/READ/fastp2CloseRef.bam.depth
```

Load data:
```{r}
#db
zymo.bed <- read.table(md5("../../Source/REF/zymo/D6305.rRNA.barrnap.bed",
                           "d322af1f013b4c037b74eedacb38f65b"),col.names = c(
 "species","from","to","unit","primary","strand","from2","to2","color"))
fungi.bed <- read.table(md5("../../Source/REF/fungi/sanger/sanger.mock7.rRNA.ITSx.bed",
                           "51c084d9add03b4f63a617e57e774938"),col.names = c(
 "species","from","to","unit","primary","strand","from2","to2","color"))

#phenotype
metadata <- read.csv("../../Results/JAN20/metadata.csv")
mockdata <- metadata%>%filter(SID%in%c("Z1","Z2","Z3"))

# previous res
BM.BI.msh.tsv <- read.table(
  md5("../../Results/JAN20/STAT/mash/B.M.BI.msh.tsv","4d0d4fdf31a9229d495b16b38ffe42f3"),
  col.names = c("SID","BID","BB","rpb","kmers"))
FM.BI.msh.tsv <- read.table(
  md5("../../Results/JAN20/STAT/mash/F.M.BI.msh.tsv","3471149b554cf2de66d86906f3680174"),
  col.names = c("SID","BID","BB","rpb","kmers"))

# current res
##bac
zymo.cov <- read.table(
  md5("../../Results/JAN20/STAT/CLIP/B.M.clip2CloseRef.ref.cov","c4f3b16a0fbb2e0a480ec25e2f529dac"),
  col.names = c("SID","species","pos","cov"))

BM.bead.anno <- read.table(
  md5("../../Results/JAN20/STAT/CLIP/B.M.clip2CloseRef.bead.anno","1c97c719ce7c18238b4da675128d1254"),
   col.names = c("SID","BID","BLen","asms","clips","tax.clips","tax.pieces","tax",
                 "ident","qMapLen","mismatch","gap","bitScore","bitScore2nd",
                 "sMapPieces","sMapLen","MAPDESC","strands","qMaps","sMaps"
))

#fungi
fungi.cov <- read.table(
  md5("../../Results/JAN20/STAT/CLIP/F.M.clip2CloseRef.ref.cov","77cae57be81c69d78770570f06eb26a0"),
  col.names = c("SID","species","pos","cov"))

FM.bead.anno <- read.table(
  md5("../../Results/JAN20/STAT/CLIP/F.M.clip2CloseRef.bead.anno","1c58a9f465c59b9ebd81f29e175a58dd"),
   col.names = c("SID","BID","BLen","asms","clips","tax.clips","tax.pieces","tax",
                 "ident","qMapLen","mismatch","gap","bitScore","bitScore2nd",
                 "sMapPieces","sMapLen","MAPDESC","strands","qMaps","sMaps"
))

#test
test.cov <- read.table("../../Results/JAN20/STAT/READ/fastp2CloseRef.bam.depth",
                       col.names = c("SID","species","pos","cov"))
```

Data curation:
```{r}
badf <- merge(metadata,rbind(BM.bead.anno,FM.bead.anno),by="SID")
badf$delta.clips <- badf$clips-badf$tax.clips
getMapPropFun <- function(d){
  uOrder <- c("ALL","SSU","ITS1","5.8S","ITS2","LSU")
  ulist <- data.frame(key=c("ALL","SSU","16S","ITS","ITS1","5.8S","ITS2","LSU","23S"),
                      ind=c(1,2,2,3,3,4,5,6,6))
  getM <- stri_match_all_regex(d,"(ALL|23S|LSU|ITS|ITS1|ITS2|5.8S|SSU|16S)\\((\\d+)%\\)")
  getN <- length(getM)
  res <- matrix(0,ncol = 6,nrow = getN)
  for(i in 1:length(getM)){
    geti <- getM[[i]]
    for(j in 1:nrow(geti)){
      res[i,ulist$ind[which(ulist$key==geti[j,2])]] <- geti[j,3]
    }
  }
  res <- matrix(as.numeric(res),ncol=6)
  colnames(res) <- uOrder
  return(res)
}
badf <- cbind(badf,getMapPropFun(badf$MAPDESC))
badf$mainUnit <- ifelse(badf$SSU>badf$LSU,"SSU","LSU")
```

```{r}
ggplot(badf,aes(x=ident,linetype=SID,color=factor(delta.clips))) + geom_density() +
  scale_x_continuous(breaks=seq(84,100,2),limits=c(85,100)) + facet_grid(kindom~.)
```

```{r}
delta.clips.stat <- ddply(badf,c("SID","kindom","delta.clips"),summarise,count=length(BID))
delta.clips.stat <- ddply(
  delta.clips.stat,c("SID"),transform,
  pct=round(count/sum(count),4),cumsum=cumsum(round(count/sum(count),4)))
ggplot(delta.clips.stat,aes(x = "x" , y = pct, fill = factor(delta.clips))) + 
  geom_bar(stat = 'identity', position = 'fill', width = 1) + coord_polar(theta = 'y') +
  facet_wrap(~SID,ncol=3) + labs(x = '', y = '', title = '') + theme_bw() +
  geom_text(aes(y = 1- cumsum + pct/2, x = 1, label = paste0(pct*100,"%"))) +
  theme(axis.text = element_blank()) + theme(axis.ticks = element_blank()) 

```

```{r}
badf$clip5 <- factor(ifelse(badf$clips>5,">5",badf$clips),levels=c(seq(1,5),">5"))
clips.stat <- ddply(badf,c("SID","clip5"),summarise,count=length(BID))
clips.stat <- ddply(
  clips.stat,c("SID"),transform,
  pct=round(count/sum(count),4),cumsum=cumsum(round(count/sum(count),4)))
ggplot(clips.stat,aes(x = "x" , y = pct, fill = clip5)) + 
  geom_bar(stat = 'identity', position = 'fill', width = 1) + coord_polar(theta = 'y') +
  facet_wrap(~SID,ncol=3) + labs(x = '', y = '', title = '') + theme_bw() +
  geom_text(aes(y = 1- cumsum + pct/2, x = 1, label = paste0(pct*100,"%"))) +
  theme(axis.text = element_blank()) + theme(axis.ticks = element_blank())

```


```{r}
ggplot(badf,aes(x=BLen,y=sMapLen)) + 
  stat_density_2d(aes(fill = stat(nlevel)), geom = "polygon") + 
  scale_fill_gradientn(colours=rev(brewer.pal(11,"Spectral"))) +
  facet_wrap(~SID,ncol=3,scale="free") + 
  ggtitle("Length of Beads and mapped reference")
```

```{r}
ggplot(badf,aes(x=sMapLen/BLen,color=factor(delta.clips))) + geom_density()+
  facet_wrap(~SID,ncol=2,dir="v",scale="free") + 
  ggtitle("Length proportion of mapped length")
```

```{r}
baMdat <- melt(badf[,c("SID","BID","ALL","SSU","ITS1","5.8S","ITS2","LSU")],
  id.vars = c("SID","BID"),variable.name = "unit",value.name = "pct")

ggplot(baMdat%>%filter(pct>0),aes(x=pct,color=unit)) + geom_density() +
  facet_wrap(~SID,ncol=3,scale="free")
```

```{r}
baMdat2 <- melt(badf[,c("SID","BID","sMapLen","ALL","SSU","ITS1","5.8S","ITS2","LSU")],
  id.vars = c("SID","BID","sMapLen"),variable.name = "unit",value.name = "pct")

ggplot(baMdat2%>%filter(pct>0), aes(x=sMapLen,y=pct)) +
  stat_density_2d(aes(fill = stat(nlevel)), geom = "polygon") + 
  scale_fill_gradientn(colours=rev(brewer.pal(11,"Spectral"))) +
  facet_grid(unit~SID,scales="free") + 
  ggtitle("Relationship of sMapLen and pct")

ggplot(baMdat2%>%filter(pct>0&sMapLen>999), aes(x=sMapLen,y=pct)) +
  stat_density_2d(aes(fill = stat(nlevel)), geom = "polygon") + 
  scale_fill_gradientn(colours=rev(brewer.pal(11,"Spectral"))) +
  facet_grid(unit~SID,scales="free") + 
  ggtitle("Relationship of sMapLen and pct")

ggplot(baMdat2%>%filter(pct>0&sMapLen>2999), aes(x=sMapLen,y=pct)) +
  stat_density_2d(aes(fill = stat(nlevel)), geom = "polygon") + 
  scale_fill_gradientn(colours=rev(brewer.pal(11,"Spectral"))) +
  facet_grid(unit~SID,scales="free") + 
  ggtitle("Relationship of sMapLen and pct")
```

CloseRef bed coverage:
```{r}
covTemplateFun <- function(d){
  d <- d%>%filter(unit!="5S_rRNA")
  d$unit <- droplevels(d$unit)
  dmin <- min(d$from)
  dmax <- max(d$to)
  res <- data.frame(species=d$species[1],pos=seq(1,dmax))
  res$unit <- factor(NA,levels=levels(d$unit))
  for(i in 1:nrow(d)){
    res$unit[d$from[i]:d$to[i]] <- d$unit[i]
  }
  res <- res[dmin:dmax,]
  return(res)
}

#bac
BM.bed.tmp <- ddply(zymo.bed,"species",covTemplateFun)
BM.bed.cov.df <- merge(zymo.cov,BM.bed.tmp,by=c("species","pos"),all=T)
BM.bed.cov.df <- BM.bed.cov.df%>%filter(species%in%levels(zymo.cov$species))
BM.bed.cov.df$cov[which(is.na(BM.bed.cov.df$cov))] <- 0
levels(BM.bed.cov.df$unit) <- c("SSU","LSU","ITS1")
#fungi
FM.bed.tmp <- ddply(fungi.bed,"species",covTemplateFun)
FM.bed.cov.df <- merge(fungi.cov,FM.bed.tmp,by=c("species","pos"),all=T)
FM.bed.cov.df <- FM.bed.cov.df%>%filter(species%in%levels(fungi.cov$species))
FM.bed.cov.df$cov[which(is.na(FM.bed.cov.df$cov))] <- 0

bed.cov.df <- rbind(cbind(kingdom="bac",BM.bed.cov.df), cbind(kingdom="fungi",FM.bed.cov.df))
#basic stat
ddply(bed.cov.df,c("kingdom","species"),summarise,
      Amin=min(pos[which(cov>99)]),Amax=max(pos[which(cov>99)]),Bmax=max(pos),
      Apct=(max(pos[which(cov>99)])-min(pos[which(cov>99)]))/(max(pos)-min(pos)),
      Lmin=min(pos[which(cov>99&unit=="LSU")]),Lmax=max(pos[which(cov>99&unit=="LSU")]),
      LBmin=min(pos[which(unit=="LSU")]),LBmax=max(pos[which(unit=="LSU")]),
      Lpct=(Lmax-Lmin)/(LBmax-LBmin))
```

```{r}
ggplot(bed.cov.df,aes(x=pos,y=cov)) + 
  geom_tile(aes(y=cov/2,width=1,height=cov,fill=unit),alpha=.5) +
  geom_line(aes(linetype=SID)) + 
  facet_wrap(~kingdom+species,ncol=4,scale="free")

ggplot(bed.cov.df%>%filter(unit%in%c("ITS1","ITS2","5.8S")),aes(x=pos,y=cov)) + 
  geom_tile(aes(y=cov/2,width=1,height=cov,fill=unit),alpha=.5) +
  geom_line(aes(linetype=SID)) + scale_fill_manual(values=brewer.pal(3,"Set1")) +
  facet_wrap(~kingdom+species,ncol=4,scale="free")
```

test.cov:
```{r}
shotgun.df <- data.frame(SID="t.shotgun",read.table("../../Results/JAN20/TEST/shotgun/read2CloseRef.bam.depth",
                         col.names = c("species","pos","cov")))
test.cov.df <- merge(BM.bed.tmp,rbind(shotgun.df,test.cov),by=c("species","pos"))
ggplot(test.cov.df%>%filter(SID=="t.shotgun"),aes(x=pos,y=cov)) + 
  geom_tile(aes(y=cov/2,width=1,height=cov,fill=unit),alpha=.5) +
  geom_line(aes(linetype=SID)) + 
  facet_wrap(~species,ncol=4,scale="free")
ggplot(test.cov.df,aes(x=pos,y=cov)) + 
  geom_tile(aes(y=cov/2,width=1,height=cov,fill=unit),alpha=.5) +
  geom_line(aes(linetype=SID)) + 
  facet_wrap(~species,ncol=4,scale="free")

ggplot(test.cov.df%>%filter(unit=="ITS_rRNA"),aes(x=pos,y=cov)) + 
  geom_tile(aes(y=cov/2,width=1,height=cov,fill=unit),alpha=.5) +
  geom_line(aes(linetype=SID)) + scale_fill_manual(values=brewer.pal(3,"Set1")[2]) +
  facet_wrap(~species,ncol=4,scale="free")
```
fixed ref cov:
```{r}
zymo.fix.bed <- read.table("../../Source/REF/zymo/D6305.fix.rRNA.barrnap.bed",col.names = c(
 "species","from","to","unit","primary","strand","from2","to2","color"))
zymo.fix.cov <- read.table("../../Results/JAN20/STAT/CLIP/clip2CloseRef.fix.ref.cov",
                           col.names = c("SID","species","pos","cov"))
BM.fix.bed.tmp <- ddply(zymo.fix.bed,"species",covTemplateFun)
BM.fix.bed.cov.df <- merge(zymo.fix.cov,BM.fix.bed.tmp,by=c("species","pos"),all=T)
BM.fix.bed.cov.df <- BM.fix.bed.cov.df%>%filter(species%in%levels(zymo.fix.cov$species))
BM.fix.bed.cov.df$cov[which(is.na(BM.fix.bed.cov.df$cov))] <- 0
levels(BM.fix.bed.cov.df$unit) <- c("SSU","LSU","ITS1")

ggplot(BM.fix.bed.cov.df,aes(x=pos,y=cov)) + 
  geom_tile(aes(y=cov/2,width=1,height=cov,fill=unit),alpha=.5) +
  geom_line(aes(linetype=SID)) + 
  facet_wrap(~species,ncol=4,scale="free")

```

Merge with kmers:
```{r}
bkadf <- merge(BI.msh.tsv[,-3],badf,by=c("SID","BID"))

ggplot(bkadf%>%filter(ALL>0), aes(x=kmers,y=ALL)) +
  stat_density_2d(aes(fill = stat(nlevel)), geom = "polygon") + 
  scale_fill_gradientn(colours=rev(brewer.pal(11,"Spectral"))) +
  facet_grid(SID~.,scales="free") + 
  ggtitle("Relationship of kmers and ALL pct")

ggplot(bkadf%>%filter(ALL>0&mainUnit=="16S"), aes(x=kmers,y=`16S`)) +
  stat_density_2d(aes(fill = stat(nlevel)), geom = "polygon") + 
  scale_fill_gradientn(colours=rev(brewer.pal(11,"Spectral"))) +
  facet_grid(SID~.,scales="free") + 
  ggtitle("Relationship of kmers and ALL pct")

ggplot(bkadf%>%filter(ALL>0&mainUnit=="23S"), aes(x=kmers,y=`23S`)) +
  stat_density_2d(aes(fill = stat(nlevel)), geom = "polygon") + 
  scale_fill_gradientn(colours=rev(brewer.pal(11,"Spectral"))) +
  facet_grid(SID~.,scales="free") + 
  ggtitle("Relationship of kmers and ALL pct")


bkadf$kmerH <- round(bkadf$kmers/200,0)*200
ggplot(bkadf%>%filter(kmerH<9999), aes(x=factor(kmerH),y=sMapLen)) + geom_boxplot() +
  facet_grid(SID~.,scales="free") + 
  ggtitle("Relationship of kmers and ALL pct") + theme(axis.text.x = element_text(angle=-90))
```

See BI with coverage > 90%
```{r}
bkadf90 <- bkadf%>%filter(ALL>80)

ggplot(bkadf90,aes(x=tax,fill=SID)) +geom_bar() + coord_flip()

ggplot(bkadf90,aes(x=kmers,linetype=SID)) +geom_density() 
ggplot(bkadf90,aes(x=ident,linetype=SID)) +geom_density() 
```


# Compare with beads distance file
```{r}
dist.Z1 <- read.table("../../Results/JAN20/SAM/Z1/CLIP/all.clip.rename.clust.debug.dist.tsv",
                      col.names = c("BID","BID2","clips","dist"))
BM.bead.anno.Z1 <- BM.bead.anno%>%filter(SID=="Z1")

dist.Z1.df <- merge(dist.Z1,BM.bead.anno.Z1[,c("BID","clips","tax","ident")],by.x="BID",by.y="BID")
dist.Z1.df <- merge(dist.Z1.df,BM.bead.anno.Z1[,c("BID","clips","tax","ident")],by.x="BID2",by.y="BID")
colnames(dist.Z1.df) <- c("BID2","BID","overlapClips","dist","clips1","tax1","ident1","clips2","tax2","ident2")
dist.Z1.df$checkTax <- ifelse(dist.Z1.df$tax1==dist.Z1.df$tax2,"Same","Diff")
```

```{r}
```

```{r}
ggplot(dist.Z1.df,aes(x=-dist/overlapClips,fill=checkTax)) +geom_histogram(position="dodge") +scale_x_log10()
```

```{r}
ggplot(dist.Z1.df, aes(x=ident1,y=ident2,color=checkTax)) +
  stat_density_2d() + 
  scale_x_continuous(breaks=seq(90,100,2)) + 
  scale_y_continuous(breaks=seq(90,100,2)) + 
  scale_fill_gradientn(colours=rev(brewer.pal(11,"Spectral")))+
  ggtitle("Relationship of sMapLen and pct")
```


```{r}
stat.dist.Z1 <- ddply(dist.Z1.df,c("BID"),summarise,
                      Scount=sum(overlapClips[checkTax=="Same"]),
                      Dcount=sum(overlapClips[checkTax=="Diff"]),
                      Sdist=sum(dist[checkTax=="Same"]),
                      Ddist=sum(dist[checkTax=="Diff"])
                      )

ggplot(stat.dist.Z1, aes(x=Scount+Dcount,y=Sdist/(Ddist+Sdist))) +
  stat_density_2d() + 
  scale_fill_gradientn(colours=rev(brewer.pal(11,"Spectral")))+
  ggtitle("Relationship of sMapLen and pct")


```

```{r}

```


















#FIN.