---
title: "JAN20 DATA ASSESSMENT"
author: "Chao"
date: "1/21/2020"
output: 
  html_notebook:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)

library(ggplot2)
library(reshape2)
library(treeio)
library(ggtree)
library(ggtreeExtra) # BiocManager::install("ggtreeExtra")

do.write<- F

calTreeLv <- function(tree){
  root_i <- rootnode(tree)
  res <- data.frame(parent=root_i,node=root_i,level=0,tip=F)
  k=1
  while(k){
    root_k <- NULL
    for( i in root_i){
      childs <- child(tree,i)
      if(length(childs) == 0){
        res$tip[which(res$node==i)] <- T
      }else{
        res <- rbind(
          res,
          data.frame(parent=i,node=childs,level=k,tip=F))
      }
      root_k <- c(root_k,childs)
    }
    if(length(root_k) == 0 ){
      break
    }
    root_i <- root_k
    k = k + 1
  }
  return(res)
}

get.sub.treedata <- function(tree_df,select_nodes){
  subtree <- tree_df%>%filter(node%in%select_nodes)
  tips <- setdiff(subtree$node,subtree$parent)
  nodes <- setdiff(subtree$parent,tips)
  newOrder <- c(sort(tips),sort(nodes))
  subtree$parent <- unlist(sapply(subtree$parent,function(x){which(newOrder==x)}))
  subtree$node <- unlist(sapply(subtree$node,function(x){which(newOrder==x)}))

  return(as.treedata(subtree))
}


  
```

# Background

Ref:\
1. <https://guangchuangyu.github.io/ggtree-book/chapter-ggtree.html>\
2. <https://yulab-smu.github.io/treedata-book/chapter2.html> 3. <https://cran.r-project.org/web/packages/tidytree/tidytree.pdf>

**Get subunits sequence**

```{bash, eval=FALSE}
cd $LFR/Assay/15.AUG20

for i in {S,O}{1,2,3};do 
  awk -v n=$i 'FNR%2==1{if(/>16S_rRNA/){gsub("16S_rRNA::",n);p=1}else{p=0}} (p==1){print}' SAM/$i/CLIP/all.barrnap.fa;
done > SAM/SBR/CLIP/all.SSU.fa
for i in {S,O}{1,2,3};do 
  awk -v n=$i 'FNR%2==1{if(/>23S_rRNA/){gsub("23S_rRNA::",n);p=1}else{p=0}} (p==1){print}' SAM/$i/CLIP/all.barrnap.fa;
done > SAM/SBR/CLIP/all.LSU.fa

for i in {S,O}{1,2,3};do
  perl -e '($tag,$f1,$f2)=@ARGV;open I,"< $f1";while(<I>){if($_=~/16S/&&$_!~/only \d+ percent of/){
      @s=split /\t/;$HS{$s[0]}=($s[3]-1)."-".$s[4]."($s[6])";}};close I;
    open I,"< $f2";$/=">";while(<I>){@s=split /\n/;@d=split /:/,$s[0];if($d[3] eq $HS{$d[2]}){print ">$tag$d[2]:$d[3]\n$s[1]\n"}};close I;' \
    $i SAM/$i/CLIP/all.barrnap.gff SAM/$i/CLIP/all.barrnap.fa
done > SAM/SBR/CLIP/full.SSU.fa

for i in {S,O}{1,2,3};do
  perl -e '($tag,$f1,$f2)=@ARGV;open I,"< $f1";while(<I>){if($_=~/23S/&&$_!~/only \d+ percent of/){@s=split /\t/;$HS{$s[0]}++}};close I;
    open I,"< $f2";$/=">";while(<I>){@s=split /\n/;@d=split /:/,$s[0];if(exists $HS{$d[2]}){print ">$tag$d[2]:$d[3]\n$s[1]\n"}};close I;' \
    $i SAM/$i/CLIP/all.barrnap.gff SAM/$i/CLIP/all.barrnap.fa
done > SAM/SBR/CLIP/full.LSU.fa

perl tmp.getValidCIDFA.pl SAM/SBR/KRAKEN/BEAD+/data/added.txt ../../Results/DEC20/STAT/KRAKEN/SXR.r2m.bracken.metadata.tsv SAM/SBR/CLIP/full.SSU.fa > SAM/SBR/CLIP/PBA.SSU.fa

perl tmp.getValidCIDFA.pl SAM/SBR/KRAKEN/BEAD+/data/added.txt ../../Results/DEC20/STAT/KRAKEN/SXR.r2m.bracken.metadata.tsv SAM/SBR/CLIP/all.LSU.fa > SAM/SBR/CLIP/PBA.LSU.fa

```

Get species level OTUs:

```{bash,eval=FALSE}
mkdir -p SAM/SBR/PREDICT/
metabbq IO tree2fa -s SAM/SBR/CLIP/PBA.SSU.fa -l SAM/SBR/CLIP/PBA.LSU.fa -r species -d CLADE0.99 -t SAM/SBR/KRAKEN/BEAD+/data/added.txt -o SAM/SBR/PREDICT/species.fa -p -v
```

    [tree2fa] start
      Reading  ... done
      Reading SAM/SBR/CLIP/PBA.SSU.fa ... done
      Reading SAM/SBR/CLIP/PBA.LSU.fa ... done
      Reading SAM/SBR/KRAKEN/BEAD+/data/added.txt ... done
      Writing to SAM/SBR/PREDICT/species.fa ...done
      Written 5590 clade seqs. 18073 records cannot found seqs.
      SSU: 5574 (279+5295+0) | LSU: 16(1+15+0) | BOTH:0(0+0+0)
      *Each unit: <total number> (<from rank> + <from clade> + <from linkage clade>
    [All done!]

**RUN RAxML** species: ssu subunits joint:

```{bash, eval=FALSE}
mafft --thread 16 SAM/SBR/PREDICT/species.SSU.fa > SAM/SBR/PREDICT/species.ssu.mafft.fa
trimal -gappyout -in SAM/SBR/PREDICT/species.ssu.mafft.fa -out SAM/SBR/PREDICT/species.ssu.mafft.trimal.fa

raxmlHPC-HYBRID-AVX -f a -p 12345 -s SAM/SBR/PREDICT/species.ssu.mafft.trimal.fa -x 12345 -T 20 -# 100 -m GTRCAT  -w $PWD/SAM/SBR/PREDICT -n species.ssu.raxml
```

**RUN RAxML-NG** species: ssu subunits joint:

```{bash, eval=FALSE}
raxml-ng --check --msa SAM/SBR/PREDICT/species.ssu.mafft.trimal.fa --model GTR+G --prefix SAM/SBR/PREDICT/T1

raxml-ng --parse --msa SAM/SBR/PREDICT/species.ssu.mafft.trimal.fa --model GTR+G --prefix SAM/SBR/PREDICT/T2

raxml-ng --msa SAM/SBR/PREDICT/species.ssu.mafft.trimal.fa --model GTR+G --prefix SAM/SBR/PREDICT/T3 --threads 40 --seed 2
```

Get genus level OTUs:

```{bash,eval=FALSE}
metabbq IO tree2fa -s SAM/SBR/CLIP/PBA.SSU.fa -l SAM/SBR/CLIP/PBA.LSU.fa -r genus -d CLADE0.97 -t SAM/SBR/KRAKEN/BEAD+/data/added.txt -o SAM/SBR/PREDICT/genus.fa -p -v
```

    [tree2fa] start
      Reading  ... done
      Reading SAM/SBR/CLIP/PBA.SSU.fa ... done
      Reading SAM/SBR/CLIP/PBA.LSU.fa ... done
      Reading SAM/SBR/KRAKEN/BEAD+/data/added.txt ... done
      Writing to SAM/SBR/PREDICT/genus.fa ...done
      Written 2793 clade seqs. 33184 records cannot found seqs.
      SSU: 2782 (228+2554+0) | LSU: 14(0+14+0) | BOTH:3(0+3+0)
      *Each unit: <total number> (<from rank> + <from clade> + <from linkage clade>
    [All done!]

**RUN RAxML** genus: both subunits joint:

```{bash, eval=FALSE}
mafft --thread 16 SAM/SBR/PREDICT/genus.SSU.fa > SAM/SBR/PREDICT/genus.ssu.mafft.fa
trimal -gappyout -in SAM/SBR/PREDICT/genus.ssu.mafft.fa -out SAM/SBR/PREDICT/genus.ssu.mafft.trimal.fa

raxmlHPC-HYBRID-AVX -f a -p 12345 -s SAM/SBR/PREDICT/genus.ssu.mafft.trimal.fa -x 12345 -T 40 -# 100 -m GTRCAT  -w $PWD/SAM/SBR/PREDICT -n genus.ssu.raxml
```

#read real data

```{r, eval=FALSE}
read.raxml2 <- function (file) 
{
  tree.text <- readLines(file)
  #tree_text <- gsub("(:[0-9\\.eE+\\-]+)\\[(\\d+)\\]", "\\@\\2\\1", tree.text)
  tree_text <- gsub("\\)(\\d+)(:[0-9\\.eE+\\-]+)", "\\)\\@\\1\\2", tree.text)
  phylo <- read.tree(text = tree_text)
  if (any(grepl("@", phylo$node.label))) {
    bootstrap <- as.numeric(gsub("[^@]*@(\\d+)", "\\1", 
      phylo$node.label))
    phylo$node.label <- gsub("@\\d+", "", phylo$node.label)
  }
  if (all(phylo$node.label == "")) {
    phylo$node.label <- NULL
  }
  bootstrap <- data_frame(node = Ntip(phylo) + 1:phylo$Nnode, bootstrap = bootstrap)
  new("treedata", file = file, treetext = tree.text,phylo = phylo, data = bootstrap)
}
phylum.raxml <- read.raxml2("../../Results/DEC20/SAM/SBR/PREDICT/RAxML_bipartitions.phylum+ref.ssu.raxml")
get.fields((phylum.raxml))

p <- ggtree(phylum.raxml) + geom_tiplab() #+ geom_label(aes(label=bootstrap))

p2 <- p %>% collapse(node=65) + 
  geom_point2(aes(subset=(node==65)), shape=21, size=5, fill='green')

p2
```

# read external info:

```{r, eval=FALSE}
# awk -F "\t" '($0!~"PUB"){print $2"\t"$3"\t"$4"\t"$6"\t"$7"\t"$1}' SAM/SBR/KRAKEN/BEAD+/data/added.txt >  SAM/SBR/KRAKEN/BEAD+/data/added.tsv

path.info <- read.table("../../Results/DEC20/SAM/SBR/KRAKEN/BEAD+/data/added.tsv",quote="",
       sep="\t",fill = NA,col.names = c("id","rank","clade","anno","SID","path"),as.is = T)


genus.raxml <- read.raxml("../../Results/DEC20/SAM/SBR/PREDICT/RAxML_bipartitionsBranchLabels.genus.ssu.raxml")
p.info <- read.table("../../Results/DEC20/SAM/SBR/PREDICT/genus.SSU.inf",sep="\t",fill=NA,
                     col.names = c("id","cladeName","SID","length","len","clades","annos"))

p.info.df <- as_tibble(merge(data.frame(label=as.phylo(genus.raxml)$tip.label),
      merge(p.info,path.info,by="id"),by.x="label",by.y="id",all.x=T))
p.info.df$cladeName <- ifelse(is.na(p.info.df$cladeName),as.character(p.info.df$label),p.info.df$cladeName)
p.info.df$anno <- ifelse(is.na(p.info.df$anno)|p.info.df$anno=="",as.character(p.info.df$cladeName),p.info.df$anno)
p.info.df$anno2 <- ifelse(grepl("CLADE",p.info.df$anno),"",p.info.df$anno)

tmp <- strsplit(p.info.df$path,";")
p.info.df$domain <- sapply(tmp,"[",1)
p.info.df$phylum <- sapply(tmp,"[",2)
p.info.df$class  <- sapply(tmp,"[",3)
p.info.df$order  <- sapply(tmp,"[",4)
p.info.df$family <- sapply(tmp,"[",5)

p.info.df$domain <- ifelse(grepl("CLADE",p.info.df$domain),"",p.info.df$domain)
p.info.df$phylum <- ifelse(grepl("CLADE",p.info.df$phylum),"",p.info.df$phylum)

p.tree <- full_join(genus.raxml, p.info.df, by="label")

p <- ggtree(p.tree,layout="circular") + geom_tiplab(aes(label=anno))

if(do.write){
ggsave(p,width=200,height=200,limitsize = FALSE,
       filename = "../../Results/DEC20/SAM/SBR/PREDICT/genus.plot.pdf")
}
```

# Get tree levels

```{r}

genus.phylo.lvs <- calTreeLv(genus.raxml)
```

# Summary annotations

```{r, eval=FALSE}
tree_tb <- as_tibble(p.tree)
tree_tb$offsprings <- 0

for(i in max(genus.phylo.lvs$level):1){
  #lv_df <- genus.phylo.lvs %>% filter(level==i) %>% group_by(parent) %>% sumNodeFun(tree_tb)
  for(p in unique((genus.phylo.lvs %>% filter(level==i))$parent)){
    ind <- which(tree_tb$node==p)
    df_i <- tree_tb %>% filter(parent == p)
    for(rank_i in c("domain","phylum","class","order", "family")){
      taxon <- df_i[,rank_i,T]
      taxon[which(is.na(taxon))] <- "unknown"
      r_tab1 <- table(unlist(lapply(strsplit(df_i[,rank_i,T],"__"),"[",1)))
      r_tab2 <- table(unlist(lapply(strsplit(df_i[,rank_i,T],"__"),"[",2)))

      if(length(r_tab1)==1){
        if(length(r_tab2)==1){
          tree_tb$rank[ind] <- names(r_tab1)[1]
          tree_tb$annos[ind] <- paste0(names(r_tab1)[1],"__",names(r_tab2)[1])
          tree_tb[ind,rank_i] <- paste0(names(r_tab1)[1],"__",names(r_tab2)[1])
        }else{
          tree_tb$rank[ind] <- names(r_tab1)[1]
          tree_tb$annos[ind] <- paste0(names(r_tab1)[1],"__Mixed")
          tree_tb[ind,rank_i] <- "Mixed"
        }
      }else{
        tree_tb$rank[ind] <- "Mixed"
        tree_tb$annos[ind] <- "Mixed_Mixed"
        tree_tb[ind,rank_i] <- "Mixed"
      }
    }
    if(is.na(tree_tb$label[ind])){
      tree_tb$label[ind] <- paste0("n",tree_tb$node[ind])
    }
    tree_tb$offsprings[ind] <- sum(df_i$offsprings) + nrow(df_i)
  }
}
```

```{r, eval=FALSE}
df <- as_tibble(data.frame(parent=c(5,5,4,4,4),node=c(1,2,3,4,5),
                            branch.length=c(1e-6,2e-6,3e-6,NA,0.001),
                           label=c("t1","t2","t3",NA,NA)),
               )
class(df) <- c("tbl_tree",class(df))

ggtree(as.phylo(df))
ape::rtree()
```

```{r, eval=FALSE}
load("test.RData")
#save.image("test.RData")
select_nodes <- genus.phylo.lvs$node[which(genus.phylo.lvs$level<=5)]


genus.subset.phylo <- get.sub.treedata(tree_tb,select_nodes)

ggtree(genus.subset.phylo)
```

```{r, eval=FALSE}
select_nodes <- genus.phylo.lvs$node[which(genus.phylo.lvs$level<=32)]

genus.subset.phylo <- as.phylo(p.tree) %>% drop.tip(., .$tip.label[-select_nodes], rooted = TRUE) %>%
  left_join(select(tree_tb,-c("parent","node","branch.length","bootstrap")),by="label")

genus.subset.df <- as_tibble(genus.subset.phylo)
phylum.freq <- table(genus.subset.df$phylum)
show.phylum.color <- names(phylum.freq[which(phylum.freq>1)])
genus.subset.df$phylum.shown <- ifelse(genus.subset.df$phylum%in%show.phylum.color,genus.subset.df$phylum,"others")

genus.tree3 <- as.treedata(genus.subset.df)
p <- ggtree(genus.tree3,aes(color=phylum.shown)) +
  geom_tiplab(aes(label=node)) + 
  geom_nodelab(aes(label=node)) + 
  theme_tree2()

p
ggtree::flip(p,726) %>% 
  ggtree::rotate(727)

ggtree::flip(p,764，727) %>% 
  ggtree::rotate(768)
```

```{r, eval=FALSE}
p2 <- ggtree(genus.tree3,aes(color=phylum.shown)) +
  geom_tiplab(aes(label=node)) + 
  theme_tree2()

ggtree::rotate(p2,726) %>% 
  ggtree::rotate(727)
```

```{r, eval=FALSE}
tdf <- genus.subset.df%>%filter(label%in%genus.tree3@phylo$tip.label)

p %<+% tdf + geom_tippoint(aes(color=phylum.shown)) +
  geom_facet(panel = "Trait", data=tdf[,c("label","offsprings")],
               geom = geom_col, 
               aes(x = offsprings.x,color=phylum.shown),
               orientation = 'y') 
```

remove clade info, only show annotated clades

```{r}
p2 <- ggtree(p.tree,layout="circular") + geom_tiplab(aes(label=phylum),align=T)
p2f <-  ggtree(p.tree,layout="circular") %>%
  collapse(node=parent(p.tree,nodeid(p.tree,15143))) %>% collapse(node=parent(p.tree,nodeid(p.tree,947611))) +
  geom_tiplab(aes(label=phylum),align=TRUE,linesize=.3,hjust=1) 
p2f
#p4 <- ggtree(p.tree) + geom_tiplab(align=TRUE,aes(label=path),linesize=.1,hjust=0,size=1)

if(do.write){
ggsave(p2f,width=200,height=200,limitsize = FALSE,
       filename = "../../Results/DEC20/SAM/SBR/PREDICT/genus.onlyAnnoPhylum.plot.pdf")
}
```

# species tree:

```{r}
sp.raxml <- read.tree("../../Results/DEC20/SAM/SBR/PREDICT/T3.raxml.bestTree")
sp.info <- read.table("../../Results/DEC20/SAM/SBR/PREDICT/species.SSU.inf",sep="\t",fill = NA,
                     col.names = c("id","cid","SID","length","len","clades","annos"))
sp.path.info <- read.table("../../Results/DEC20/SAM/SBR/KRAKEN/BEAD+/data/added.tsv",
       sep="\t",fill = NA,col.names = c("id","rank","clade","anno","SID","path"),as.is = T)

sp.info.df <- as_tibble(merge(data.frame(label=as.phylo(sp.raxml)$tip.label),
      merge(sp.info,sp.path.info,by="id"),by.x="label",by.y="id",all.x=T))

sp.info.df$clade <- ifelse(is.na(sp.info.df$clade),as.character(sp.info.df$label),sp.info.df$clade)
sp.info.df$anno <- ifelse(is.na(sp.info.df$anno)|sp.info.df$anno=="",as.character(sp.info.df$clade),sp.info.df$anno)
sp.info.df$anno2 <- ifelse(grepl("CLADE",sp.info.df$anno),"",sp.info.df$anno)

tmp <- strsplit(sp.info.df$path,";")
sp.info.df$domain <- ifelse(grepl(".*domain__(\\w+);.*",sp.info.df$path,perl=T),
                            sub(".*(domain__\\w+);.*","\\1",sp.info.df$path, perl=T),NA)
sp.info.df$phylum <- ifelse(grepl(".*;phylum__(\\w+);.*",sp.info.df$path,perl=T),
                            sub(".*;(phylum__\\w+);.*","\\1",sp.info.df$path, perl=T),NA)
sp.info.df$class  <- ifelse(grepl(".*;class__(\\w+);.*",sp.info.df$path,perl=T),
                            sub(".*;(class__\\w+);.*","\\1",sp.info.df$path, perl=T),NA)
sp.info.df$order  <- ifelse(grepl(".*;order__(\\w+);.*",sp.info.df$path,perl=T),
                            sub(".*(order__\\w+);.*","\\1",sp.info.df$path, perl=T),NA)
sp.info.df$family <- ifelse(grepl(".*;family__(\\w+);.*",sp.info.df$path,perl=T),
                            sub(".*(family__\\w+);.*","\\1",sp.info.df$path, perl=T),NA)
sp.info.df$genus <- ifelse(grepl(".*;genus__(\\w+);.*",sp.info.df$path,perl=T),
                            sub(".*(genus__\\w+);.*","\\1",sp.info.df$path, perl=T),NA)
sp.info.df$species <- ifelse(grepl(".*;species__(\\w+);.*",sp.info.df$path,perl=T),
                            sub(".*(species__\\w+);.*","\\1",sp.info.df$path, perl=T),NA)

sp.info.df$domain <- ifelse(grepl("CLADE",sp.info.df$domain),"",sp.info.df$domain)
sp.info.df$phylum <- ifelse(grepl("CLADE",sp.info.df$phylum),"",sp.info.df$phylum)
sp.info.df$class <- ifelse(grepl("CLADE",sp.info.df$class),"",sp.info.df$class)
sp.info.df$order <- ifelse(grepl("CLADE",sp.info.df$order),"",sp.info.df$order)
sp.info.df$family <- ifelse(grepl("CLADE",sp.info.df$family),"",sp.info.df$family)
sp.info.df$genus <- ifelse(grepl("CLADE",sp.info.df$genus),"",sp.info.df$genus)
sp.info.df$species <- ifelse(grepl("CLADE",sp.info.df$species),"",sp.info.df$species)


sp.tree <- full_join(sp.raxml, sp.info.df, by="label")


```

## Summary annotations

```{r}
species.phylo.lvs <- calTreeLv(sp.tree)

tree_tb <- as_tibble(sp.tree)
tree_tb$offsprings <- 0

for(i in max(species.phylo.lvs$level):1){
  #lv_df <- species.phylo.lvs %>% filter(level==i) %>% group_by(parent) %>% sumNodeFun(tree_tb)
  for(p in unique((species.phylo.lvs %>% filter(level==i))$parent)){
    ind <- which(tree_tb$node==p)
    df_i <- tree_tb %>% filter(parent == p)
    for(rank_i in c("domain","phylum","class","order", "family","genus","species")){
      taxon <- df_i[,rank_i,T]
      taxon[which(is.na(taxon))] <- "__"
      r_tab1 <- table(unlist(lapply(strsplit(df_i[,rank_i,T],"__"),"[",1)))
      r_tab2 <- table(unlist(lapply(strsplit(df_i[,rank_i,T],"__"),"[",2)))

      if(length(r_tab1)==1){
        if(length(r_tab2)==1){
          tree_tb$rank[ind] <- names(r_tab1)[1]
          tree_tb$annos[ind] <- paste0(names(r_tab1)[1],"__",names(r_tab2)[1])
          tree_tb[ind,rank_i] <- paste0(names(r_tab1)[1],"__",names(r_tab2)[1])
        }else{
          tree_tb$rank[ind] <- names(r_tab1)[1]
          tree_tb$annos[ind] <- paste0(names(r_tab1)[1],"__Mixed")
          tree_tb[ind,rank_i] <- "Mixed"
        }
      }else{
        tree_tb$rank[ind] <- "Mixed"
        tree_tb$annos[ind] <- "Mixed_Mixed"
        tree_tb[ind,rank_i] <- "Mixed"
      }
    }
    if(is.na(tree_tb$label[ind])){
      tree_tb$label[ind] <- paste0("n",tree_tb$node[ind])
    }
    tree_tb$offsprings[ind] <- sum(df_i$offsprings) + nrow(df_i)
  }
}

```

## tree with all sp. level

```{r}
#save.image(file="SBR.sp.tree.RData")
#load("SBR.sp.tree.RData")
phylum.freq <- table(tree_tb$phylum)
#Bdellovibrionota
#Patescibacteria
show.phylum.color <- c(names(rev(sort(phylum.freq))[1:12]),"Patescibacteria")
tree_tb$phylum.shown <- ifelse(
  is.na(tree_tb$phylum)|tree_tb$phylum == "","Unknown",
  ifelse(tree_tb$phylum%in%show.phylum.color,tree_tb$phylum,"Others"))
tree_tb$phylum.shown <- factor(tree_tb$phylum.shown,
                               levels=c(show.phylum.color,"Others","Unknown"))

#
tree_tb.tr <- tree_tb[,c("parent","node","branch.length","label","phylum")]
tree_tb.df <- tree_tb[,c("node","label","phylum.shown")]
### 

p0f <- ggtree(as.treedata(tree_tb.tr),layout="fan",) +
  geom_tiplab(align=TRUE,aes(label=phylum)) +
  geom_nodelab(aes(label=node))

p0f <- ggtree(as.treedata(tree_tb.tr)) +
  geom_tiplab(align=T,aes(label=phylum)) +
  geom_nodelab(aes(label=node))

p1f <- flip(p0f, 7753,8608) %>% rotate(9741) +
  geom_hilight(data=tree_tb.df%>%filter(
    !phylum.shown%in%c("",NA,"Mixed","others")),
    mapping=aes(node=node,fill=phylum.shown),alpha=0.1,size=0.05)
if(do.write){
  ggsave(p1f,width=30,height=200,limitsize = FALSE,
       filename = "../../Results/DEC20/SAM/SBR/PREDICT/species.ver.onlyAnnoPhylum.with.node.plot.pdf")

}

### 180

p2f0 <- ggtree(as.treedata(tree_tb.tr),layout="circular")

p2f1 <- open_tree(p2f0,180) %>% rotate_tree(90) %>% flip(7753,8608) %>% 
  rotate(9741) %>% rotate(7640) + 
  geom_nodelab(aes(label=node)) +
  geom_hilight(data=tree_tb.df%>%filter(!phylum.shown%in%c("Mixed")),
               mapping=aes(node=node,fill=phylum.shown), alpha=0.1, size=0.05)

if(do.write){
  ggsave(p2f1,width=200,height=200,limitsize = FALSE,
       filename = "../../Results/DEC20/SAM/SBR/PREDICT/species.180.onlyAnnoPhylum.with.node.plot.pdf")

}

p2f2 <- open_tree(p2f0,180) %>% rotate_tree(90) %>% flip(7753,8608) %>% rotate(9741) %>% rotate(7640) +
  geom_hilight(data=tree_tb.df%>%filter(!phylum.shown%in%c("Mixed")),mapping=aes(node=node,fill=phylum.shown), alpha=0.1, size=0.05)

if(do.write){
  ggsave(p2f2,width=200,height=200,limitsize = FALSE,
       filename = "../../Results/DEC20/SAM/SBR/PREDICT/species.180.onlyAnnoPhylum.without.node.plot.pdf")

}
```

## Summary nodes info


```{r}
getCladeDf <- function(treedata,nodes){
  clade.df <- NULL
  clade.nodes <- NULL
  for(i in nodes){
    i.df <- offspring(treedata,i)
    clade.nodes <- c(clade.nodes,i.df$node)
  }
  clade.nodes <- unique(clade.nodes)
  clade.df <- treedata%>%filter(node%in%clade.nodes)
  clade.tips <- setdiff(clade.df$node,clade.df$parent)
  clade.df <- clade.df%>%filter(node%in%clade.tips)

  clade.table <- table(clade.df$phylum)
  clade.pie <- pie(clade.table)
  return(list(df=clade.df,tb=clade.table,pie=clade.pie))
}




Acidobacteriota.obj <- getCladeDf(tree_tb,c(5304,5612,5630,5684,5845,5847))
others2.obj <- getCladeDf(tree_tb,c(5907))
Verrucomicrobiota.obj <- getCladeDf(tree_tb,c(6025))
Planctomycetota.obj <- getCladeDf(tree_tb,c(6124))
Bacteroidota5.obj <- getCladeDf(tree_tb,c(6619))
Patescibacteria.obj <- getCladeDf(tree_tb,c(6352))
Chloroflexi.obj <- getCladeDf(tree_tb,c(6371))
Firmicutes.obj <- getCladeDf(tree_tb,c(6642))
Actinobacteriota.obj <- getCladeDf(tree_tb,c(6793))
Others10.obj <- getCladeDf(tree_tb,c(7496,7686,7503,7647))
Myxococcota.obj <- getCladeDf(tree_tb,c(7522,7641,7676))
Gemmatimonadota.obj <- getCladeDf(tree_tb,9842)
Proteobacteria.obj <- getCladeDf(tree_tb,
  c(9939,9952,9752,9827,9959,9961,10536,10118,7715,8608,7754))
Bdellovibrionota.obj <-  getCladeDf(tree_tb,8575)
Bacteroidota15.obj <- getCladeDf(tree_tb,7766)




SBR.summary.clades.df <- rbind(
  cbind(Acidobacteriota.obj$df,area="1"),
  cbind(others2.obj$df,area="2"),
  cbind(Verrucomicrobiota.obj$df,area="3"),
  cbind(Planctomycetota.obj$df,area="4"),
  cbind(Bacteroidota5.obj$df,area="5"),
  cbind(Patescibacteria.obj$df,area="6"),
  cbind(Chloroflexi.obj$df,area="7"),
  cbind(Firmicutes.obj$df,area="8"),
  cbind(Actinobacteriota.obj$df,area="9"),
  cbind(Others10.obj$df,area="10"),
  cbind(Myxococcota.obj$df,area="11"),
  cbind(Gemmatimonadota.obj$df,area="12"),
  cbind(Proteobacteria.obj$df,area="13"),
  cbind(Bdellovibrionota.obj$df,area="14"),
  cbind(Bacteroidota15.obj$df,area="15")
)
SBR.summary.clades.df$area <- factor(SBR.summary.clades.df$area,
                                     levels=seq(1,15))
```

```{r}
SBR.summary.clades.df$area <- factor(SBR.summary.clades.df$area,
                                     levels=seq(1,15))

ggplot(SBR.summary.clades.df,aes(x=factor(area))) + 
  geom_bar(aes(fill=phylum.shown),position="fill") + 
  coord_flip() + theme_bw()
```

```{r}
sumFun <- function(d){
  freq <- rev(sort(table(d$phylum)))
  top1 <- sub("phylum__","",names(freq)[1])
  count <- freq[1]
  unk <- length(which(d$phylum=="Unknown"|is.na(d$phylum)))
  text <- paste(count,unk,nrow(d),sep="/")
  res <- data.frame(topTax=top1,c.top=count,unk=unk,total=nrow(d),text=text)
  return(res)
}
summary.df <- ddply(SBR.summary.clades.df,"area",sumFun)
summary.df
```


```{r}
summary.df[,c(2,6)]
```

```{r}
SBR.summary.clades.df$is.phylum <- ifelse(is.na(SBR.summary.clades.df$phylum)|SBR.summary.clades.df$phylum=="",
                                          "Unknown","Known") 
SBR.summary.clades.df$is.class <- ifelse(is.na(SBR.summary.clades.df$class)|SBR.summary.clades.df$class=="",
                                         "Unknown","Known") 
SBR.summary.clades.df$is.order <- ifelse(is.na(SBR.summary.clades.df$order)|SBR.summary.clades.df$order=="",
                                         "Unknown","Known") 
SBR.summary.clades.df$is.family <- ifelse(is.na(SBR.summary.clades.df$family)|SBR.summary.clades.df$family=="",
                                          "Unknown","Known") 
SBR.summary.clades.df$is.genus <- ifelse(is.na(SBR.summary.clades.df$genus)|SBR.summary.clades.df$genus=="",
                                         "Unknown","Known") 
SBR.summary.clades.df$is.species <- ifelse(is.na(SBR.summary.clades.df$species)|SBR.summary.clades.df$species=="",
                                           "Unknown","Known") 

SBR.unkown.stat <- melt(
  SBR.summary.clades.df, id.vars = c("SID.x","node","area"),
  measure.vars = c("is.phylum","is.class","is.order","is.family","is.genus","is.species"),
  variable.name = "Rank",value.name = "annotation")


SBR.unkown.stat$Annotation <- factor(SBR.unkown.stat$annotation,levels=c("Unknown","Known"))
SBR.unkown.stat$Rank <- as.factor(SBR.unkown.stat$Rank)
levels(SBR.unkown.stat$Rank) <- c("P","C","O","F","G","S")
ggplot(SBR.unkown.stat,aes(x=Rank)) + 
  geom_bar(stat="count",position="fill",aes(fill=Annotation),color="black") +
  facet_grid(area~.) + theme_bw() + 
  scale_fill_manual(values = c("grey90","deepskyblue3")) +
  scale_y_continuous(breaks=seq(0,1,1))

if(do.write){
  ggsave(width=4,height=6,limitsize = FALSE,
       filename = "../../Results/DEC20/SAM/SBR/PREDICT/SBR.knownAnno.stat.barplot.pdf")

}
```


#FIN.
