---
title: "JAN20 DATA ASSESSMENT"
author: "Chao"
date: "1/21/2020"
output: 
  html_notebook:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(treeio)
library(ggtree)
library(ggplot2)
```

# Background

Ref:  
1. https://guangchuangyu.github.io/ggtree-book/chapter-ggtree.html  
2. https://yulab-smu.github.io/treedata-book/chapter2.html
3. https://cran.r-project.org/web/packages/tidytree/tidytree.pdf

**Get subunits sequence**
```{bash, eval=FALSE}
cd $LFR/Assay/15.AUG20

for i in {S,O}{1,2,3};do 
  awk -v n=$i 'FNR%2==1{if(/>16S_rRNA/){gsub("16S_rRNA::",n);p=1}else{p=0}} (p==1){print}' SAM/$i/CLIP/all.barrnap.fa;
done > SAM/SBR/CLIP/all.SSU.fa
for i in {S,O}{1,2,3};do 
  awk -v n=$i 'FNR%2==1{if(/>23S_rRNA/){gsub("23S_rRNA::",n);p=1}else{p=0}} (p==1){print}' SAM/$i/CLIP/all.barrnap.fa;
done > SAM/SBR/CLIP/all.LSU.fa

for i in {S,O}{1,2,3};do
  perl -e '($tag,$f1,$f2)=@ARGV;open I,"< $f1";while(<I>){if($_=~/16S/&&$_!~/only \d+ percent of/){
      @s=split /\t/;$HS{$s[0]}=($s[3]-1)."-".$s[4]."($s[6])";}};close I;
    open I,"< $f2";$/=">";while(<I>){@s=split /\n/;@d=split /:/,$s[0];if($d[3] eq $HS{$d[2]}){print ">$tag$d[2]:$d[3]\n$s[1]\n"}};close I;' \
    $i SAM/$i/CLIP/all.barrnap.gff SAM/$i/CLIP/all.barrnap.fa
done > SAM/SBR/CLIP/full.SSU.fa

for i in {S,O}{1,2,3};do
  perl -e '($tag,$f1,$f2)=@ARGV;open I,"< $f1";while(<I>){if($_=~/23S/&&$_!~/only \d+ percent of/){@s=split /\t/;$HS{$s[0]}++}};close I;
    open I,"< $f2";$/=">";while(<I>){@s=split /\n/;@d=split /:/,$s[0];if(exists $HS{$d[2]}){print ">$tag$d[2]:$d[3]\n$s[1]\n"}};close I;' \
    $i SAM/$i/CLIP/all.barrnap.gff SAM/$i/CLIP/all.barrnap.fa
done > SAM/SBR/CLIP/full.LSU.fa

perl tmp.getValidCIDFA.pl SAM/SBR/KRAKEN/BEAD+/data/added.txt ../../Results/DEC20/STAT/KRAKEN/SXR.r2m.bracken.metadata.tsv SAM/SBR/CLIP/full.SSU.fa > SAM/SBR/CLIP/PBA.SSU.fa

perl tmp.getValidCIDFA.pl SAM/SBR/KRAKEN/BEAD+/data/added.txt ../../Results/DEC20/STAT/KRAKEN/SXR.r2m.bracken.metadata.tsv SAM/SBR/CLIP/all.LSU.fa > SAM/SBR/CLIP/PBA.LSU.fa

```

Get species level OTUs:
```{bash,eval=FALSE}
mkdir -p SAM/SBR/PREDICT/
metabbq IO tree2fa -s SAM/SBR/CLIP/PBA.SSU.fa -l SAM/SBR/CLIP/PBA.LSU.fa -r species -d CLADE0.99 -t SAM/SBR/KRAKEN/BEAD+/data/added.txt -o SAM/SBR/PREDICT/species.fa -p -v
```
```
[tree2fa] start
  Reading  ... done
  Reading SAM/SBR/CLIP/PBA.SSU.fa ... done
  Reading SAM/SBR/CLIP/PBA.LSU.fa ... done
  Reading SAM/SBR/KRAKEN/BEAD+/data/added.txt ... done
  Writing to SAM/SBR/PREDICT/species.fa ...done
  Written 5590 clade seqs. 18073 records cannot found seqs.
  SSU: 5574 (279+5295+0) | LSU: 16(1+15+0) | BOTH:0(0+0+0)
  *Each unit: <total number> (<from rank> + <from clade> + <from linkage clade>
[All done!]
```
**RUN RAxML **
species: ssu subunits joint:
```{bash, eval=FALSE}
mafft --thread 16 SAM/SBR/PREDICT/species.SSU.fa > SAM/SBR/PREDICT/species.ssu.mafft.fa
trimal -gappyout -in SAM/SBR/PREDICT/species.ssu.mafft.fa -out SAM/SBR/PREDICT/species.ssu.mafft.trimal.fa

raxmlHPC-HYBRID-AVX -f a -p 12345 -s SAM/SBR/PREDICT/species.ssu.mafft.trimal.fa -x 12345 -T 20 -# 100 -m GTRCAT  -w $PWD/SAM/SBR/PREDICT -n species.ssu.raxml
```

**RUN RAxML-NG **
species: ssu subunits joint:
```{bash, eval=FALSE}
raxml-ng --check --msa SAM/SBR/PREDICT/species.ssu.mafft.trimal.fa --model GTR+G --prefix SAM/SBR/PREDICT/T1

raxml-ng --parse --msa SAM/SBR/PREDICT/species.ssu.mafft.trimal.fa --model GTR+G --prefix SAM/SBR/PREDICT/T2

raxml-ng --msa SAM/SBR/PREDICT/species.ssu.mafft.trimal.fa --model GTR+G --prefix SAM/SBR/PREDICT/T3 --threads 40 --seed 2
```

Get genus level OTUs:
```{bash,eval=FALSE}
metabbq IO tree2fa -s SAM/SBR/CLIP/PBA.SSU.fa -l SAM/SBR/CLIP/PBA.LSU.fa -r genus -d CLADE0.97 -t SAM/SBR/KRAKEN/BEAD+/data/added.txt -o SAM/SBR/PREDICT/genus.fa -p -v
```
```
[tree2fa] start
  Reading  ... done
  Reading SAM/SBR/CLIP/PBA.SSU.fa ... done
  Reading SAM/SBR/CLIP/PBA.LSU.fa ... done
  Reading SAM/SBR/KRAKEN/BEAD+/data/added.txt ... done
  Writing to SAM/SBR/PREDICT/genus.fa ...done
  Written 2793 clade seqs. 33184 records cannot found seqs.
  SSU: 2782 (228+2554+0) | LSU: 14(0+14+0) | BOTH:3(0+3+0)
  *Each unit: <total number> (<from rank> + <from clade> + <from linkage clade>
[All done!]
```

**RUN RAxML **
genus: both subunits joint:
```{bash, eval=FALSE}
mafft --thread 16 SAM/SBR/PREDICT/genus.SSU.fa > SAM/SBR/PREDICT/genus.ssu.mafft.fa
trimal -gappyout -in SAM/SBR/PREDICT/genus.ssu.mafft.fa -out SAM/SBR/PREDICT/genus.ssu.mafft.trimal.fa

raxmlHPC-HYBRID-AVX -f a -p 12345 -s SAM/SBR/PREDICT/genus.ssu.mafft.trimal.fa -x 12345 -T 40 -# 100 -m GTRCAT  -w $PWD/SAM/SBR/PREDICT -n genus.ssu.raxml
```


#read real data
```{r}
read.raxml2 <- function (file) 
{
  tree.text <- readLines(file)
  #tree_text <- gsub("(:[0-9\\.eE+\\-]+)\\[(\\d+)\\]", "\\@\\2\\1", tree.text)
  tree_text <- gsub("\\)(\\d+)(:[0-9\\.eE+\\-]+)", "\\)\\@\\1\\2", tree.text)
  phylo <- read.tree(text = tree_text)
  if (any(grepl("@", phylo$node.label))) {
    bootstrap <- as.numeric(gsub("[^@]*@(\\d+)", "\\1", 
      phylo$node.label))
    phylo$node.label <- gsub("@\\d+", "", phylo$node.label)
  }
  if (all(phylo$node.label == "")) {
    phylo$node.label <- NULL
  }
  bootstrap <- data_frame(node = Ntip(phylo) + 1:phylo$Nnode, bootstrap = bootstrap)
  new("treedata", file = file, treetext = tree.text,phylo = phylo, data = bootstrap)
}
phylum.raxml <- read.raxml2("../../Results/DEC20/SAM/SBR/PREDICT/RAxML_bipartitions.phylum+ref.ssu.raxml")
get.fields((phylum.raxml))

p <- ggtree(phylum.raxml) + geom_tiplab() #+ geom_label(aes(label=bootstrap))

p2 <- p %>% collapse(node=65) + 
  geom_point2(aes(subset=(node==65)), shape=21, size=5, fill='green')

p2
```

# read external info:
```{r}
# awk -F "\t" '($0!~"PUB"){print $2"\t"$3"\t"$4"\t"$6"\t"$7"\t"$1}' SAM/SBR/KRAKEN/BEAD+/data/added.txt >  SAM/SBR/KRAKEN/BEAD+/data/added.tsv

path.info <- read.table("../../Results/DEC20/SAM/SBR/KRAKEN/BEAD+/data/added.tsv",
       sep="\t",fill = NA,col.names = c("id","rank","clade","anno","SID","path"),as.is = T)


genus.raxml <- read.raxml("../../Results/DEC20/SAM/SBR/PREDICT/RAxML_bipartitionsBranchLabels.genus.ssu.raxml")
p.info <- read.table("../../Results/DEC20/SAM/SBR/PREDICT/genus.SSU.inf",sep="\t",fill=NA,
                     col.names = c("id","cladeName","SID","length","len","clades","annos"))

p.info.df <- as_tibble(merge(data.frame(label=as.phylo(genus.raxml)$tip.label),
      merge(p.info,path.info,by="id"),by.x="label",by.y="id",all.x=T))
p.info.df$cladeName <- ifelse(is.na(p.info.df$cladeName),as.character(p.info.df$label),p.info.df$cladeName)
p.info.df$anno <- ifelse(is.na(p.info.df$anno)|p.info.df$anno=="",as.character(p.info.df$cladeName),p.info.df$anno)
p.info.df$anno2 <- ifelse(grepl("CLADE",p.info.df$anno),"",p.info.df$anno)

tmp <- strsplit(p.info.df$path,";")
p.info.df$domain <- sapply(tmp,"[",1)
p.info.df$phylum <- sapply(tmp,"[",2)
p.info.df$class  <- sapply(tmp,"[",3)
p.info.df$order  <- sapply(tmp,"[",4)
p.info.df$family <- sapply(tmp,"[",5)

p.info.df$domain <- ifelse(grepl("CLADE",p.info.df$domain),"",p.info.df$domain)
p.info.df$phylum <- ifelse(grepl("CLADE",p.info.df$phylum),"",p.info.df$phylum)

p.tree <- full_join(genus.raxml, p.info.df, by="label")

p <- ggtree(p.tree,layout="circular") + geom_tiplab(aes(label=anno))
p

if(do.write){
ggsave(p,width=200,height=200,limitsize = FALSE,
       filename = "../../Results/DEC20/SAM/SBR/PREDICT/genus.plot.pdf")
}
```
# Get tree levels 
```{r}
calTreeLv <- function(tree){
  root_i <- rootnode(tree)
  res <- data.frame(parent=root_i,node=root_i,level=0,tip=F)
  k=1
  while(k){
    root_k <- NULL
    for( i in root_i){
      childs <- child(tree,i)
      if(length(childs) == 0){
        res$tip[which(res$node==i)] <- T
      }else{
        res <- rbind(
          res,
          data.frame(parent=i,node=childs,level=k,tip=F))
      }
      root_k <- c(root_k,childs)
    }
    if(length(root_k) == 0 ){
      break
    }
    root_i <- root_k
    k = k + 1
  }
  return(res)
}

genus.phylo.lvs <- calTreeLv(genus.raxml)
```


# Summary annotations
```{r}
tree_tb <- as_tibble(p.tree)
tree_tb$offsprings <- 0

for(i in max(genus.phylo.lvs$level):1){
  #lv_df <- genus.phylo.lvs %>% filter(level==i) %>% group_by(parent) %>% sumNodeFun(tree_tb)
  for(p in unique((genus.phylo.lvs %>% filter(level==i))$parent)){
    ind <- which(tree_tb$node==p)
    df_i <- tree_tb %>% filter(parent == p)
    for(rank_i in c("domain","phylum","class","order", "family")){
      taxon <- df_i[,rank_i,T]
      taxon[which(is.na(taxon))] <- "unknown"
      r_tab1 <- table(unlist(lapply(strsplit(df_i[,rank_i,T],"__"),"[",1)))
      r_tab2 <- table(unlist(lapply(strsplit(df_i[,rank_i,T],"__"),"[",2)))

      if(length(r_tab1)==1){
        if(length(r_tab2)==1){
          tree_tb$rank[ind] <- names(r_tab1)[1]
          tree_tb$annos[ind] <- paste0(names(r_tab1)[1],"__",names(r_tab2)[1])
          tree_tb[ind,rank_i] <- paste0(names(r_tab1)[1],"__",names(r_tab2)[1])
        }else{
          tree_tb$rank[ind] <- names(r_tab1)[1]
          tree_tb$annos[ind] <- paste0(names(r_tab1)[1],"__Mixed")
          tree_tb[ind,rank_i] <- "Mixed"
        }
      }else{
        tree_tb$rank[ind] <- "Mixed"
        tree_tb$annos[ind] <- "Mixed_Mixed"
        tree_tb[ind,rank_i] <- "Mixed"
      }
    }
    if(is.na(tree_tb$label[ind])){
      tree_tb$label[ind] <- paste0("n",tree_tb$node[ind])
    }
    tree_tb$offsprings[ind] <- sum(df_i$offsprings) + nrow(df_i)
  }
}
```

```{r}
df <- as_tibble(data.frame(parent=c(5,5,4,4,4),node=c(1,2,3,4,5),
                            branch.length=c(1e-6,2e-6,3e-6,NA,0.001),
                           label=c("t1","t2","t3",NA,NA)),
               )
class(df) <- c("tbl_tree",class(df))

ggtree(as.phylo(df))
ape::rtree()
```

```{r}
load("test.RData")
#save.image("test.RData")
select_nodes <- genus.phylo.lvs$node[which(genus.phylo.lvs$level<=2)]

subtree <- tree_tb%>%filter(node%in%select_nodes)
  
genus.subset.phylo <- as.phylo(subtree)
genus.subset.phylo$edge[,1] <- unlist(sapply(genus.subset.phylo$edge[,1],function(x){which(sort(select_nodes)==x)}))
genus.subset.phylo$edge[,2] <- unlist(sapply(genus.subset.phylo$edge[,2],function(x){which(sort(select_nodes)==x)}))

nodeMap <- subtree$node
subtree <- as.data.frame(subtree)
subtree$parent <- unlist(lapply(subtree$parent,function(x){which(nodeMap==x)}))
subtree$node <- unlist(lapply(subtree$node,function(x){which(nodeMap==x)}))
class(subtree) <- c("tbl_tree","tbl","data.frame")
genus.subset.phylo <- as.phylo(subtree)

ggtree(genus.subset.phylo)
```

```{r}
select_nodes <- genus.phylo.lvs$node[which(genus.phylo.lvs$level<=20)]

genus.subset.phylo <- as.phylo(p.tree) %>% drop.tip(., .$tip.label[-select_nodes], rooted = TRUE) %>%
  left_join(select(tree_tb,-c("parent","node","branch.length","bootstrap")),by="label")

genus.subset.df <- as_tibble(genus.subset.phylo)
phylum.freq <- table(genus.subset.df$phylum)
show.phylum.color <- names(phylum.freq[which(phylum.freq>1)])
genus.subset.df$phylum.shown <- ifelse(genus.subset.df$phylum%in%show.phylum.color,genus.subset.df$phylum,"others")

genus.tree3 <- as.treedata(genus.subset.df)
p <- ggtree(genus.tree3,aes(color=phylum.shown)) +
  geom_tiplab(aes(label=node)) + 
  geom_nodelab(aes(label=node)) + 
  theme_tree2()

p
ggtree::rotate(p,766) %>% 
  ggtree::rotate(727)
```

```{r}
tdf <- genus.subset.df%>%filter(label%in%genus.tree3@phylo$tip.label)

p %<+% tdf + geom_tippoint(aes(color=phylum.shown)) +
  geom_facet(panel = "Trait", data=tdf[,c("label","offsprings")],
               geom = geom_col, 
               aes(x = offsprings.x,color=phylum.shown),
               orientation = 'y') 
```


remove clade info, only show annotated clades
```{r}
p2 <- ggtree(p.tree,layout="circular") + geom_tiplab(aes(label=phylum),align=T)
p2f <-  ggtree(p.tree,layout="circular") %>%
  collapse(node=parent(p.tree,nodeid(p.tree,15143))) %>% collapse(node=parent(p.tree,nodeid(p.tree,947611))) +
  geom_tiplab(aes(label=phylum),align=TRUE,linesize=.3,hjust=1) 
p2f
#p4 <- ggtree(p.tree) + geom_tiplab(align=TRUE,aes(label=path),linesize=.1,hjust=0,size=1)

if(do.write){
ggsave(p2f,width=200,height=200,limitsize = FALSE,
       filename = "../../Results/DEC20/SAM/SBR/PREDICT/genus.onlyAnnoPhylum.plot.pdf")
}
```


#test
```{r}

```

# test species tree:
```{r}
sp.raxml <- read.raxml2("../../Results/JAN20/SAM/SBR/PREDICT/RAxML_bipartitions.species+ref.ssu.raxml")
sp.info <- read.table("../../Results/JAN20/SAM/SBR/PREDICT/species.SSU.inf",sep="\t",fill = NA,
                     col.names = c("id","SID","length","len","clades","annos"))

sp.info.df <- as_tibble(merge(data.frame(label=as.phylo(sp.raxml)$tip.label),
      merge(sp.info,path.info,by="id"),by.x="label",by.y="id",all.x=T))
sp.info.df$clade <- ifelse(is.na(sp.info.df$clade),as.character(sp.info.df$label),sp.info.df$clade)
sp.info.df$anno <- ifelse(is.na(sp.info.df$anno)|sp.info.df$anno=="",as.character(sp.info.df$clade),sp.info.df$anno)
sp.tree <- full_join(sp.raxml, sp.info.df, by="label")

sp <- ggtree(sp.tree,layout="circular") #+ geom_tiplab(aes(label=anno))
sp

p2 <- sp %>% collapse(node=671) + 
  geom_point2(aes(subset=(node==671)), shape=21, size=5, fill='green')

p2


```



# treeio practices:
```{r}
beast_file <- system.file("examples/MCC_FluA_H3.tree", 
                          package="ggtree")
rst_file <- system.file("examples/rst", package="ggtree")
mlc_file <- system.file("examples/mlc", package="ggtree")

beast_tree <- read.beast(beast_file)
codeml_tree <- read.codeml(rst_file, mlc_file)
merged_tree <- merge_tree(beast_tree, codeml_tree)

get.fields(merged_tree)
```
```{r}
library(dplyr)
df <- fortify(merged_tree) %>%
  select(dN_vs_dS, dN, dS, rate) %>%
  subset(dN_vs_dS >=0 & dN_vs_dS <= 1.5) %>%
  tidyr::gather(type, value, dN_vs_dS:dS)
df$type[df$type == 'dN_vs_dS'] <- 'dN/dS'
df$type <- factor(df$type, levels=c("dN/dS", "dN", "dS"))
ggplot(df, aes(rate, value)) + geom_hex() + 
  facet_wrap(~type, scale='free_y') 
```


Read from Raxml:
```{r}
raxml_file <- system.file("extdata/RAxML/",
                          "RAxML_bipartitionsBranchLabels.H3", 
                          package="treeio")
raxml <- read.raxml(raxml_file)
get.fields((raxml))
ggtree(raxml)

```
Time-scaled layout:
```{r}
beast_file <- system.file("examples/MCC_FluA_H3.tree", 
                          package="ggtree")
beast_tree <- read.beast(beast_file)
ggtree(beast_tree, mrsd="2013-01-01",layout="circular") + theme_tree2()
```

Annotation of clade
```{r}
treetext = "(((ADH2:0.1[&&NHX:S=human], ADH1:0.11[&&NHX:S=human]):
0.05 [&&NHX:S=primates:D=Y:B=100],ADHY:
0.1[&&NHX:S=nematode],ADHX:0.12 [&&NHX:S=insect]):
0.1[&&NHX:S=metazoa:D=N],(ADH4:0.09[&&NHX:S=yeast],
ADH3:0.13[&&NHX:S=yeast], ADH2:0.12[&&NHX:S=yeast],
ADH1:0.11[&&NHX:S=yeast]):0.1[&&NHX:S=Fungi])[&&NHX:D=N];"
tree <- read.nhx(textConnection(treetext))
ggtree(tree) + geom_tiplab(align=TRUE,linesize=.3,hjust=1) + 
  geom_label(aes(x=branch, label=S), fill='lightgreen') + 
  geom_label(aes(label=D), fill='steelblue') + 
  geom_text(aes(label=B), hjust=-.5,vjust=-1)
```
```{r}
ggtree(tree,layout="circular",aes(color=factor(label))) + geom_tiplab(align=TRUE,linesize=NA,hjust=1) + 
  geom_label(aes(x=branch, label=S), fill='lightgreen') + 
  geom_label(aes(label=D), fill='steelblue') + 
  geom_text(aes(label=B), hjust=-.5,vjust=-1)
```


```{r}
nwk <- system.file("extdata", "sample.nwk", package="treeio")
tree <- read.tree(nwk)
p <- ggtree(tree) + geom_tiplab(align=TRUE)
viewClade(p, MRCA(p, "I", "L"))

```

```{r}
p2 <- p %>% collapse(node=21) + 
  geom_point2(aes(subset=(node==21)), shape=21, size=5, fill='green')
p2 <- collapse(p2, node=23) + 
  geom_point2(aes(subset=(node==23)), shape=23, size=5, fill='red')
print(p2)
expand(p2, node=23) %>% expand(node=21)
```



























#FIN.